%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

%\vspace{-0.35in}
This work would not be possible without the tireless commitment and mentoring by my friend and supervisor, Dr. Stefan Reinsberg.
Stefan, I could not have done this without you - you made my journey everything that it was.
Dr. Baker was also instrumental in helping run, plan, and provide support for all of the work presented here - thank you Jenn! 
Members of my supervisory committee, Drs. Kozlowski, Laule, and Minchinton were all excellent sources of advice, feedback, guidance, and general project direction. 
Particular thanks to Dr. Kozlowski for supporting and organizing weekly lab meetings giving me and other trainees to give and get honest feedback on preliminary work - I learned a tonne through these meetings.

I am fortunate enough to have some truly excellent mentors that have guided me through my PhD program and introduced me to the world of educational leadership.
Special kudos to Dr. Simon Bates who has been, and continues to be an invaluable mentor and supporter of my pursuits in teaching and learning.
Everything I know about teaching large classes has come from Simon - truly, thank you.
Both he and Dr. Ido Roll were instrumental in my initial dabbling in physics education research and Ido took me under his wing and set me up with a solid foundation in the scholarship of teaching and learning (SoTL).
I then met a community of like-minded individuals that I truly connected with on a deep and foundational level - SoTL specialists, thank you for the debates, conversations, and most importantly, your friendship.

My first big SoTL project was with Drs. James Charbonneau and Chris Addison and it dealt with the idea of measuring interdisciplinarity in undergraduate students, particularly in the Science One program.
All together we learned so much in this project and I will forever be thankful for the opportunity to work on such a fun project!
Dr. Georg Rieger gave me my first taste of teaching a class, and I will never forget the immense amount of trust and faith you placed in me.
Many former graduate students were instrumental in my growth and development, and I owe a lot to Natasha Holmes, Ellen Schelew, Amanda Parker, Mike Sitwell, Emily Altiere, Jonathan Massey, and Jared Stang - you folks were amazing and I truly look up to you!
Past, current, and former members of PHASER, thank you for giving me a place to feel at home discussing everything related to physics education.

Finally, last but certainly not least, a huge note of gratitude goes to my parents, brothers, family, and of course my wife Hira, who has supported and humoured me through my meanderings over the course of my PhD. 
\emph{I love you all.}

